### 2025-01-13 report/report_template.qmd
folders

---
title: "Group assignment: Is High-Dimensional Analysis Worth it?"
format: 
  pdf:
    toc: false
    number-sections: true
    documentclass: article
    geometry: "margin=4cm"
    bibliography: bibliography.bib
abstract: |
    You can write the abstract at the end. Do it approximately like this and you have a nice flow: one sentence about data context. One sentence about research problem. One sentence about research question. One sentence about methods compared. One sentence about results. One sentence about conclusion.
author: 
  - Carlos Perengano
  - Erika Mustermann
  - Jan Kowalski
---

# Introduction

::: {.callout-note}
Max 700 words
:::

Provide a description of your data and relevant context on the research problem it might have been designed to solve in 2 to 3 paragraphs. Maybe referring to a figure in a manner such as @fig-introimage.

```{r}
#| label: fig-introimage
#| fig-cap: A boxplot comparing the mpg of cars for different numbers of cylinders.
#| out-width: "80%"
#| echo: false
boxplot(mpg ~ cyl, data = mtcars)
```

Write max 2 paragraphs describing the research problem and leading to the methodological research question. The question should revolve around comparing high-dimensional vs not high-dimensional methods.

# Methods

::: {.callout-note}
500-600 words
:::

Description of methods you use in maximum 3 paragraphs. You might want to refer to literature which you can find in the file `bibliography.bib` e.g., @hastie2015statistical or @james2013introduction. Use bibtex notation in the `.bib` file.

Also describe your comparison or validation strategy in 1 or 2 paragraphs. Answer the question: how are you going to compare your chosen methods against each other? How are you going to conclude which is better for this dataset / research purpose?

# Results

::: {.callout-note}
Max 500 words
:::

Here, provide a description of __relevant__ results in 2 to 3 paragraphs. You are invited to use images / plots to help communicate your results, but keep this section legible and stick to the main points! In your results section, you should refrain from giving too much discussion / explanation; be quite factual.

# Conclusion

::: {.callout-note}
Max 700 words
:::

Start with a short summary of what you have done in 1 paragraph ("in this report, we have...").

Then, write your conclusion in which you answer your research question in 1 paragraph ("For this dataset, we have found that the high-dimensional analysis improved / did not improve over ..." )

Then, contextualize your conclusion by providing some discussion additional interpretation in approx. 3 paragraphs. Feel free to refer to additional literature if you like. Suggestions for future research might be nice to add.

# References
<!-- Don't put anything here, your references will be inserted automatically at the end -->